(tentative) Rotational Spectral Unmixing of Exoplanets

### 2016-10-06 paper/main.tex
minor updates

The time variability of the contribution factor $f^{\ast }$ due to spin rotation is related to the surface inhomogeneity along the equator. Approximately, $f^{\ast }_{ik}$ may be written as the weighted summation of the area fraction of $k$-th surface type in each of (the finite number $L$ of) longitudinal slices, i.e.,
%%%
\begin{eqnarray}
f^{\ast }_{ik} &=& \frac{ \int_{{\rm IV}_{i}} f_k (\vec \Omega ) \cos \theta_0 ({\vec \Omega}) \cdot \cos \theta_1 ({\vec \Omega}) \cdot d\vec \Omega }{ \int_{{\rm IV}_i}  \cos \theta_0 ({\vec \Omega}) \cdot \cos \theta_1 ({\vec \Omega}) \cdot d\vec \Omega }  \\
&=& \sum_l \frac{ \int_{{\rm IV}_{il}}  f_k (\vec \Omega ) W (\vec \Omega  ) \cdot d\vec \Omega }{ \int_{{\rm IV}}  W (\vec \Omega ) \cdot d\vec \Omega } \\
&\sim & \sum_l f_{lk} \frac{ \int_{{\rm IV}_{il}} W (\vec \Omega ) \cdot d\vec \Omega }{ \int_{{\rm IV}}  W (\vec \Omega ) \cdot d\vec \Omega } \label{eq:discretize}\\
&\equiv & \sum_l  W_{il} f_{lk} \label{eq:Wf}
\end{eqnarray}
%%%
where $l$ is the index for longitudinal slices, $f_{lk}$ is the area fraction of the $k$-th surface type in the $l$-th longitudinal slice. 
The approximation is valid only when $f_k(\vec \Omega)$ does not change or changes little in the $l$-th slice for all $k$ and $l$. 
In the last equation, $W_{il}$ is the weight function for $i$-th epoch and $l$-th longitudinal slice which depends only on the observational geometry. 
As a result,
%%%
\begin{equation}
d_{ij} = R_p^2 \sum _{l,k} W_{il} \, f_{lk} \, s_{kj}
\end{equation}
%%%
with the following condition similar to equation (\ref{eq:cond_f_ast}):
%%%
\begin{eqnarray}
\begin{cases}
\;\; 0 < s_{kj} < 1 \;\;\; & \mbox{for any $k$, $j$} \\
\;\; 0 < f_{lk} < 1 \;\;\; & \mbox{for any $l$, $k$} \\
\;\; \sum_k f_{lk} = 1 & \mbox{for any $l$} 
\end{cases}
\label{eq:constraints}
\end{eqnarray}
%%%
Now, the relevant problem is, given $d$, estimate $\{f, s\}$ subject to the constraints of equation (\ref{eq:constraints})---this is where \citet{Cowan2013} stood.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Fit with $f$ or with $f^{\ast}$?}
\label{ss:degeneracy}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\memoYF{I am not 100\% sure if the following statement is right. Need to check.} 
We argue that estimating $\{ f, s \}$ is no superior to estimating $\{ f^{\ast }, s \}$. 
One reason comes from the fact the approximation adopted in equation (\ref{eq:discretize}). This is valid only when $f_k (\vec \Omega)$ does not change significantly in the $l$-th slice. 
If this isn't the case, equation (\ref{eq:Wf}) is no longer a good discritization.

If that condition is satisfied for $L=I$ slices, then in general cases there is a one-to-one relationship between $f^{\ast}_{ik}$ and $f_{lk}$ through $W_{il}$ as far as ${\rm rank}(W)=L(=I)$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Degeneracy}
\label{ss:degeneracy}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

The time variability of the contribution factor $f^{\ast }$ due to spin rotation is related to the surface inhomogeneity along the equator. Approximately, $f^{\ast }_{ik}$ may be written as the weighted summation of the area fraction of $k$-th surface type in each of (the finite number $L$ of) longitudinal slices, i.e.,
%%%
\begin{eqnarray}
f^{\ast }_{ik} &=& \frac{ \int_{{\rm IV}_{i}} f_k (\vec \Omega ) \cos \theta_0 ({\vec \Omega}) \cdot \cos \theta_1 ({\vec \Omega}) \cdot d\vec \Omega }{ \int_{{\rm IV}_i}  \cos \theta_0 ({\vec \Omega}) \cdot \cos \theta_1 ({\vec \Omega}) \cdot d\vec \Omega }  \\
&=& \sum_l \frac{ \int_{{\rm IV}_{il}}  f_k (\vec \Omega ) W (\vec \Omega  ) \cdot d\vec \Omega }{ \int_{{\rm IV}}  W (\vec \Omega ) \cdot d\vec \Omega } \\
&\sim & \sum_l f_{lk} \frac{ \int_{{\rm IV}_{il}} W (\vec \Omega ) \cdot d\vec \Omega }{ \int_{{\rm IV}}  W (\vec \Omega ) \cdot d\vec \Omega } \label{eq:discretize}\\
&\equiv & \sum_l  W_{il} f_{lk} \label{eq:Wf}
\end{eqnarray}
%%%
where $l$ is the index for longitudinal slices, $f_{lk}$ is the area fraction of the $k$-th surface type in the $l$-th longitudinal slice. 
The approximation is valid only when $f_k(\vec \Omega)$ does not change or changes little in the $l$-th slice for all $k$ and $l$. 
In the last equation, $W_{il}$ is the weight function for $i$-th epoch and $l$-th longitudinal slice which depends only on the observational geometry. 
As a result,
%%%
\begin{equation}
d_{ij} = R_p^2 \sum _{l,k} W_{il} \, f_{lk} \, s_{kj}
\end{equation}
%%%
with the following condition similar to equation (\ref{eq:cond_f_ast}):
%%%
\begin{eqnarray}
\begin{cases}
\;\; 0 < s_{kj} < 1 \;\;\; & \mbox{for any $k$, $j$} \\
\;\; 0 < f_{lk} < 1 \;\;\; & \mbox{for any $l$, $k$} \label{eq:cond_f} \\
\;\; \sum_k f_{lk} = 1 & \mbox{for any $l$} 
\end{cases}
\label{eq:constraints}
\end{eqnarray}
%%%
Now, the relevant problem is, given $d$, estimate $\{f, s\}$ subject to the constraints of equation (\ref{eq:constraints})---this is where \citet{Cowan2013} stood.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Fit data with $f$ or with $f^{\ast}$?}
\label{ss:degeneracy}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\memoYF{I am not 100\% sure if the following statement is right. Need to check.}

We argue that estimating $\{ f, s \}$ is no superior to estimating $\{ f^{\ast }, s \}$. 
%
One reason comes from the fact the approximation adopted in equation (\ref{eq:discretize}) is valid only when $f_k (\vec \Omega)$ does not change significantly in the $l$-th slice. 
Otherwise, fitting data with $f$ by equation (\ref{eq:Wf}) includes some errors by design.

%Suppose the above approximation is justified. 
In essence, $f^{\ast }_{ik}$ and $f_{lk}$ contains the same information, where they are related through a fixed $W_{il}$ (eq. \ref{eq:Wf}). 
%
If $L=I$, in general there is a one-to-one relationship between $f^{\ast}_{ik}$ and $f_{lk}$ through $W_{il}$ as far as ${\rm rank}(W)=L(=I)$. 
In this case, clearly, equations (\ref{eq:cond_f_ast}) and (\ref{eq:cond_f}) are equivalent given $W_{il}>0$ (for any $i$ and $j$), and $\sum _l W_{il} = 1$. 
%
If $L<I$, $W$ works as a low-pass filter, as discussed in \citet{Cowan2013}. 
In this case, the inverse problem described by equation (\ref{eq:Wf}), $f^{\ast }=Wf$ is an overdetermined system, and one must to find the solution for $f$ by some means (e.g. least squares method), and the resultant $f$ may not automatically satisfy the condition listed by eqution (\ref{eq:cond_f}). 
However, this is because of the design of the problem and the assumed lowered resolution, and not because of the wrong estimation for $f$ (or $s$).

Thus, although both $\{ f^{\ast }, s \}$ and $\{ f, s \}$ could be a fitting parameters to the light curves, the estimation for $s$ should not be affected by this conversion in fitting parameters. 
In that sense, $\{ f, s \}$ is not likely to return a better solution than $\{ f^{\ast }, s \}$ does.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Strong Degeneracy}
\label{ss:degeneracy}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
